Managing variability in software systems

### 2021-05-02 Paper.tex
merge conflict

This is the approach which is used in the original FODA paper, which was extended with the introduction of cardinalities for alternative features. \cite{b10} This expansion grants the ability to create minimum and maximum constraints for choosing alternative features from a feature set, which helps with creating complex models.\\\\
In Fig. 1, one can see a small feature diagram of a car, it consists of one root node, that describes the domain we are modeling, in this case the car. The previous elaborated mandatory features are the engine and seats. Furthermore there is the option to install a cruise control and the option for a turbocharger these features are completely independet and optional. Last but not least one can choose between normal leather seats or alcantara ones but not both. Which this simplified model of a car we have all three types of features modeled

\textbf{Decision modeling}\\
asdfasdf

@@

Further explanation of these binding times can be found in the subsequent paragraphs:\\
\begin{itemize}
	\item \textit{Product architecture derivation.} The architecture of the product already binds to a variation. This is the lowest binding time and will influence all subsequent steps.\\
	\item \textit{Compilation.} When the source code is compiled there exist several options to bind to a variation by adjusting constants, macros or compiler directives.\\
	\item \textit{Linking.} This option does not exist for any technology, but for many programming languages the linking process can be manually changed. For example by switching the DLL of a module.\\
	\item \textit{Startup} A simple option to bind variants to the system is by configuration files, that influence the features a software system exhibits.\\
	\item \textit{Adaptation to runtime environment} This typically refers to plugins of a software, which can add variation to a system and can either be provided by the original developer or even third party vendors. \\
\end{itemize}
Any mechanism to bind to none, one or many different variants tend to fall into different categories presented by \cite{b2} and \cite{b3}, and also have parallels to the feature types presented in the previous section:\\
\begin{itemize}
	\item \textit{Variant Entity.} There exist several options but only one may be bound to at any given time.\\
	\item \textit{Optional Entity.} As the name suggests, an optional entity may or may not be included in the system.\\
	\item \textit{Multiple Coexisting Entities.} Similar to the variant entity but with the exception that multiple variants can be used.\\
\end{itemize}
These categories tend to be the same for the different binding times, but of course vary greatly in their implementation. While an optional entity during compilation may only be influenced by a compiler flag, it might be a whole plugin from a third party vendor when adapting a given runtime environment. An in depth analysis of these mechanisms can be found in \cite{b2}.
\section{Discussion/Challenges}
Discuss why the variability of systems is hard to tackle and why it is not always successful.
(1/4 page) JR

>>>>

This is the approach which is used in the original FODA paper, which was extended with the introduction of cardinalities for alternative features.\cite{b10} This expansion grants the ability to create minimum and maximum constraints for choosing alternative features from a feature set, which helps with creating complex models.\\\\
In Fig. 1, one can see a small feature diagram of a car, it consists of one root node, that describes the domain we are modeling, in this case the car. The previous elaborated mandatory features are the engine and seats. Furthermore there is the option to install a cruise control and the option for a turbocharger these features are completely independet and optional. Last but not least one can choose between normal leather seats or alcantara ones but not both. Which this simplified model of a car we have all three types of features modeled.\\\\
\textbf{Decision modeling}\\
In decision modeling the variability is described with decision instead of features. A variable feature in feature modeling is equal to a decision and vice versa. Contrary to feature modeling decision modeling is unable to describe the commonalities of systems.\\\\
Decision models can be represented in different ways, broadly used approaches are flat lists or tables of decisions which also includes their dependicies. In decision modeling hierarchy is a secondary aspect, because there are no uniform rules to describe hierarchy. \cite{b6}\\\\
Same as feature modeling, decision modeling has been expanded over time to enable the application coverage of DM be almost the same as FM.\cite{b6}\\\\

@@

Further explanation of these binding times can be found in the subsequent paragraphs:\\
\begin{itemize}
	\item \textit{Product architecture derivation.} The architecture of the product already binds to a variation. This is the lowest binding time and will influence all subsequent steps.\\
	\item \textit{Compilation.} When the source code is compiled there exist several options to bind to a variation by adjusting constants, macros or compiler directives.\\
	\item \textit{Linking.} This option does not exist for any technology, but for many programming languages the linking process can be manually changed. For example by switching the DLL of a module.\\
	\item \textit{Startup} A simple option to bind variants to the system is by configuration files, that influence the features a software system exhibits.\\
	\item \textit{Adaptation to runtime environment} This typically refers to plugins of a software, which can add variation to a system and can either be provided by the original developer or even third party vendors. \\
\end{itemize}
Any mechanism to bind to none, one or many different variants tend to fall into different categories presented by \cite{b2} and \cite{b3}, and also have parallels to the feature types presented in the previous section:\\
\begin{itemize}
	\item \textit{Variant Entity.} There exist several options but only one may be bound to at any given time.\\
	\item \textit{Optional Entity.} As the name suggests, an optional entity may or may not be included in the system.\\
	\item \textit{Multiple Coexisting Entities.} Similar to the variant entity but with the exception that multiple variants can be used.\\
\end{itemize}
These categories tend to be the same for the different binding times, but of course vary greatly in their implementation. While an optional entity during compilation may only be influenced by a compiler flag, it might be a whole plugin from a third party vendor when adapting a given runtime environment. An in depth analysis of these mechanisms can be found in \cite{b2}.

\section{Discussion/Challenges (1/4 page) JR}
The research field of variability had a increasing trend of published paper per year from 2000 until 2008. After that it stabilized between 25-30 new research paper per year from 2008 until 2011. One explanation of this rise in interest in the scientific community could be that the usage of variability was becoming more broadly in dynamic systems. \cite{b4}\\\\
Another aspect is that feature modeling is in its core a simple concept. Even though its impact in the industry is pretty big considering it is the most popular concept for variability modeling in software systesm until today.\cite{b9} One can say that because it is such a simple concept and due to the big amount of research in this field we understand the subject pretty well by now.\\\\
The currently existing methods also have their limitations. It is difficult to guarantee correctness and consistency, authors have to urge for verification and consistency checking. Performance overheads, runtime upgradability and the lack of support regarding security needs are some areas where the described methods are limited. Sometimes the handling of variability can create poor user-friendliness and therefore decrease usebility of a system. \cite{b4}\\\\
There are ways to decrease the negative impact of these aspects e.g. correctness and consistency can be checked by various methods such as modeling guidelines customized for specific needs, together with a reviewing process where changes to the model are checked by another person to ensure the quality of the metioned parameters.
